Prefix-free graphs and suffix array construction in sublinear space

### 2023-06-21 manuscript/main.tex
some more progress

\author[1]{Andrej Baláž}[
orcid=0000-0001-0002-0003,
email=andrejbalaz001@gmail.com,
]
\cormark[1]
\address[1]{
    Faculty of Mathematics, Physics and Informatics,
    Comenius University, Mlynská dolina F1, 842 48 Bratislava, Slovakia
}

% Footnotes
\cortext[1]{Corresponding author.}

@@

There are five cases we need to consider.
\begin{itemize}
    \item suffix represents a separator or the sentinel
    \item suffix represents a trigger or it's suffix
    \item suffix longer than the trigger representing a single position in the pangenome
    \item suffix longer than the trigger representing several positions in the pangenome
\end{itemize}

We can skip the first $n+1$ rows of the suffix table, because these suffixes
do not yield any suffix array entry.

@@

Rpair: rescaling repair with rsync \cite{2019gagie}
Prefix-free parsing for building big BWTs \cite{2019boucher}
MONI and r-index \cite{}
Prefix-free Wheeler graphs \cite{2022pfwg}
MARIA \cite{2022maria}

>>>>

\author[1]{Andrej Baláž}[
    orcid=0000-0001-0002-0003,
    email=andrejbalaz001@gmail.com,
]
\cormark[1]

\author[1]{Alessia Petescia}[
    orcid=0000,
    email=alessia@gmail.com,
]

\address[1]{
    Department of Applied Informatics,
    Faculty of Mathematics, Physics and Informatics,
    Comenius University, Bratislava, Slovakia
}

% Footnotes
\cortext[1]{Corresponding author}

@@

There are three cases where a letter at position $SA[i]$ can occur in these suffix segments:
\begin{itemize}
    \item it is a separator or the sentinel
    \item it is at the end of a segment inside the trigger part
    \item it is part of the segment outside the trigger part, separator and sentinel
\end{itemize}

In the first case, there is no corresponding position in the original pangenome,
therefore we can skip these rows of the suffix table.
% We can skip the first $n+1$ rows of the suffix table, because these suffixes
% do not yield any suffix array entry.

In the second case, the position is inside the trigger string.
Since the possitions inside the triggers are represented twice, once at the end
of a segment a second time at the beginning of the following segment, we need
to choose when to report them.
A natural choice is to report positions at the beginning of a segment and ignore
positions at the end.
Here we note that this choice also plays nicely with the previous choice of sentinel
characters, since ignoring last $k$ characters will result in not printing
sentinels either.
In conclusion, if the length of a current segment suffix is smaller or equal to
the size of the trigger words, we skip the row similarly as in the previous case.

Finaly, in the last case we report the positions.
The last case will partition the suffix table into the blocks of the same suffixes.
As an example consider the rows 20 and 21, which together for a single block.
All other blocks in the example consist of single rows and therefore we call them singletons.

This partitioning leads to three cases (from simplest to hardest):
\begin{itemize}
    \item A singleton block, occuring only once in the whole pangenome
    \item A singleton block, repeating several times in the pangenome
    \item A non-singleton block
\end{itemize}

@@

Rpair: rescaling repair with rsync \cite{2019gagie}
Prefix-free parsing for building big BWTs \cite{2019boucher}
MONI and r-index \cite{rossi2022moni,gagie2020fully}
Prefix-free Wheeler graphs \cite{2022pfwg}
MARIA \cite{2022maria}
